High angular resolution dust density and temperature maps from \Herschel\ imaging data

### 2012-08-04 documents/herschel_dust.tex
text and equations about the model

\newcommand{\foreign}[1]{\textit{#1}}
\newcommand{\etal}{\foreign{et\,al.}}
\newcommand{\documentname}{\textsl{Note}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\TheTractor}{\project{The~Tractor}}
\newcommand{\sdss}{\project{SDSS}}
\newcommand{\Herschel}{\project{Herschel}}

@@

\title{High angular resolution dust density and temperature maps from \Herschel\ imaging data}
\author{Lang, Hogg, etc}
\altaffiltext{1}{To whom correspondence should be addressed; \texttt{foo@bar.edu}}
\altaffiltext{2}{New York University}
\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie}
\altaffiltext{4}{Princeton University Observatory}

@@

This paper is a stub.  You can edit it and help \TheTractor, so long
as you are either Hogg or Lang!

\acknowledgements
It is a pleasure to thank...

>>>>

\newcommand{\foreign}[1]{\textit{#1}}
\newcommand{\etal}{\foreign{et\,al.}}
\newcommand{\documentname}{\textsl{Note}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\TheTractor}{\project{The~Tractor}}
\newcommand{\Herschel}{\project{Herschel}}
\newcommand{\PACS}{\project{PACS}}
\newcommand{\SPIRE}{\project{SPIRE}}
\newcommand{\PHAT}{\project{PHAT}}

@@

\title{High angular resolution dust density and temperature maps from \Herschel\ imaging data}
\author{
  David~W.~Hogg\altaffilmark{1,2,3},
  Dustin~Lang\altaffilmark{4,5},
  Brent~Groves\altaffilmark{3},
  Hans-Walter~Rix\altaffilmark{3},
  Karin~Sandstrom\altaffilmark{3}}
\altaffiltext{1}{To whom correspondence should be addressed; \texttt{foo@bar.edu}}
\altaffiltext{2}{Center for Cosmology and Particle Physics, New York University}
\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie}
\altaffiltext{4}{Princeton University Observatory}
\altaffiltext{5}{Carnegie Mellon University}

@@

How do you obtain reliable multi-wavelength information about a source
or a set of sources when your imaging data span a huge range of
angular scales?

\section{model}

The model consists of a \emph{physical emission model}, a
\emph{spatial model}, a set of \emph{priors} and a \emph{likelihood
  function}.  The physical emission model we choose is a toy model for
emitting dust grains with a range of temperatures and sizes.  This toy
model is that a tiny unresolved blob of dust $\Omega_k$ at temperature
$T_k$ will emit flux density $F_{\nu,k}$ (energy per unit time per
unit area per unit frequency)
\begin{eqnarray}\displaystyle
F_{\nu,k} &=& \Omega_k\,B_{\nu}(T_k)\,\left[\frac{\lambda}{\lambda_0}\right]^{-\beta_k}
\quad,
\end{eqnarray}
where $B_{\nu}(T)$ is the black-body intensity (energy per unit time
per unit area per unit solid angle per unit frequency), and $\beta_k$
is an emission parameter accounting for the radiative and
multi-temperature properties of the dust.  Usually the emission
parameter $\beta$ is around 2.  In this formulation, the dust
amplitude $\Omega_k$ has units of solid angle, and ought to be related
(perhaps not trivially) to the literal solid angle subtended by the
radiation-absorbing and emitting dust particles.  This model is
obviously a toy or effective model; much more sophisticated models
exist (CITATIONS).

The spatial model we choose is a regular grid (on a small
tangent-plane projection of the the celestial sphere) of
delta-function pixels.  This is effectively a non-parametric model (in
the sense that it has an enormous number of parameters).  To meet the
goals of this project, this grid is made as fine (in an angular sense)
as the pixel grid of the highest-resolution \Herschel\ \PACS\ images.
At each two-dimensional pixel location $\theta_k$ on the sky, there is
a dust solid angle $\Omega_k$, a temperature $T_k$, and an emission
parameter $\beta_k$.  For any real image taken at some effective
wavelength $\lambda$ there is an effective point-spread function
$\psi(\Delta\theta\given\lambda)$ with which the grid of emitting
pixels is convolved.  Thus the model intensity for an image is
\begin{eqnarray}\displaystyle
I_{\nu}(\theta,\lambda) &=& \sum_{k=1}^{K}
F_{\nu,k}\,\psi(\theta-\theta_k\given\lambda) \quad ,
\end{eqnarray}
where the convolution is trivial because the unconvolved intensity
model is a grid of delta functions.

The likelihood function---the probability of the data given the model parameters---is just the Gaussian
likelihood function
\begin{eqnarray}\displaystyle
-2\,\ln L &\approx& \chi^2
\\
\chi^2 &\equiv& \sum_{n=1}^N \frac{\left[d_n - I_{\nu}(\theta_n,\lambda_n)\right]^2}{\sigma_n^2}
\quad,
\end{eqnarray}
where the ``$\approx$'' symbol is used as a reminder that this
function is not properly normalized, each pixel $n$ has a data value
$d_n$ that is assumed to be a properly calibrated intensity, and each
pixel is at a known sky location $\theta_n$, taken in a bandpass at
effective wavelengths $\lambda_n$, and with an independent Gaussian
noise variance $\sigma_n^2$.

In non-parametric models like this---even when we only want point
estimates or optima of an objective function---we cannot avoid using
prior information to regularize the fit.  The priors we choose are
``smoothness'' priors inspired by Gaussian processes.  They penalize
the sum of squares of differences between neighboring pixels.
\begin{eqnarray}\displaystyle
-2\,\ln p(\Omega_k) &\approx& \sum_{\mathrm{neighbors}~k'} \epsilon_\Omega\,\left[\Omega_k - \Omega_{k'}\right]^2
\\
-2\,\ln p(T_k) &\approx& \sum_{\mathrm{neighbors}~k'} \epsilon_T\,\left[T_k - T_{k'}\right]^2
\\
-2\,\ln p(\beta_k) &\approx& \epsilon_2\,\left[\beta - 2\right]^2
                    + \sum_{\mathrm{neighbors}~k'} \epsilon_\beta\,\left[\beta_k - \beta_{k'}\right]^2
\\
\alpha &\equiv& \left\{\epsilon_\Omega, \epsilon_T, \epsilon_2, \epsilon_\beta\right\}
\quad ,
\end{eqnarray}
where again the ``$\approx$'' symbols are used as a reminder that
these priors are not properly normalized as written, the sum over
``neighbors'' $k'$ is a sum, for each pixel, over its four nearest
neighbors, the prior ``strength'' or information is set by the list
$\alpha$ of hyper-parameters $\{\epsilon_\Omega, \epsilon_T,
\epsilon_2, \epsilon_\beta\}$, and the emission parameter grid
$\beta_k$ is given the additional prior that we expect values around
2.

...The hyperparameters are set by cross-validation...

\section{experiments and results}

...(in detail, sky)...

...comparison to \PHAT\ extinction maps...

\section{discussion}

...Successes.

...Limitations of the emission model.

...Limitations of the spatial model, especially the PSFs.

...Limitations of the priors.

\acknowledgements It is a pleasure to thank Thomas Henning (MPIA) and
Ben Weiner (Arizona).  This project made use of the NASA
\project{Astrophysics Data System}, the Python \project{numpy},
\project{scipy}, and \project{matplotlib} open-source software
projects, and the \project{Astrometry.net} codebase.  All the code and
documentation used in this project is available as part of
\TheTractor\ at \url{http://TheTractor.org/}.
